\large{\bf Large effects and the infinitesimal model }

### 2024-01-23 review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript,
``Large effects and the infinitesimal model''.
We have addressed the remaining points of the reviwers,
hopefully to everyone's satisfaction.

@@

\vskip 4em

\pagebreak

%%%%%%%%%%%%%%
\section*{AE comments:}

\begin{quote}
As you will read from the reviewers' comments below, both of them find that the quality and clarity of the exposition has improved a lot, and they only have a list of minor (nonetheless important) comments that should be addressed in a minor revision. Provided the points raised are suitably addressed, the paper will then be accepted for publication. And though this has nothing to do with the paper, happy New Year to the authors ;-).
\end{quote}

Happy New Year!

@@

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{point}{}
     As far as I understand, the crucial assumption is that the effect sizes $\eta_{l}^{i}$
     of the alleles carried by an individual $i$ are independent
     (as in the first line of Proposition \ref{prop:parental_contribs}),
     which basically amounts to assuming linkage equilibrium in the population.
     In that case, of course, all trait values across any group of individuals are multi-variate Gaussian
     (provided effect size distributions are not heavy-tailed),
     without needing to condition on any ancestors.
     Moreover, under these conditions,
     even the trait value distribution in the population would be Gaussian
     (provided effect size distributions is not heavy-tailed).
     As far as I understand, \citet{barton2017infinitesimal} is more general
     in that it does not require $\eta_{l}^{i}$ to be independent,
     except in the founder $t=0$ generation
     (and I am not sure how crucial even that assumption is);
     in later generations, there can be an arbitrary amount of LD.
     This is a crucial point, since in general, under the infinitesimal model,
     non-Gaussian population distributions only arise because of LD, in which case,
     the $\eta_{l}^{i}$ are not independent.
     Thus, it is worth highlighting that assuming $\eta_{l}^{i}$ to be independent
     basically implies linkage equilibrium, and that this is in general,
     not a requirement for the infinitesimal model
     or for the derivations of \citet{barton2017infinitesimal},
     whereas it is a key requirement here.
     I think this had been a big part of my earlier confusion.
\end{point}

\reply{
    It is true that if the $\eta$ are not independent then $Z_M$ and $V_M$ are not independent,
    and so the Kac-Bernstein theorem does not apply.
    However -- returning to the goal of this section here -- it seems hard and maybe impossible
    to use LD to obtain a inifinitesimal-like, non-Gaussian model.
    But, we also don't see why it's impossible, and have added a note to this effect \revref.
    (Also we'd like to point out that the independence only applies to the \emph{parental} generation;
    there is in fact LD in the offspring's generation, which is the generation Proposition \ref{prop:parental_contribs}
    refers to. Our results could be extended to further generations, as we remark just before \revref,
    but since our point is about non-existence of other models, this seemed superfluous.
    Note also that although we assume independence, we don't assume the $\eta$ are identically distributed,
    so that a Gaussian or other stable distribution is not guaranteed.)
}

\begin{point}{}
     I think this also needs to be emphasised in later sections,
     e.g., section 4, where the statement that no assumption is made about the initial distribution of the population
     doesn't seem to be quite right.
     I think the implicit assumption of Linkage equilibrium does come in
     once one assumes that all the $\eta_{l}^{i}$ are independent.
\end{point}

\reply{
    We don't think this applies, as
    Section 4 is explicitly not working with the model of Section 3 --
    there are no loci explicitly modeled at all in Section 4.
    We have added a reminder to emphasize this fact \revref.
}

@@

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{point}{\revref}
     p.4, line 3 (from top): garbled sentence
\end{point}

\reply{
    We've simplified the sentence.
}

\begin{point}{\revref}
     delete one `that'
\end{point}

\reply{
    Done.
}

\begin{point}{}
     Section 6: it is unfortunate to use $\mu$ for the mutation probability and $\mu(x)$ for the death rate. Maybe use $u$ for the mutation prob.
\end{point}

\reply{
    Good call; done.
}

\begin{point}{\revref}
     p.18, line 460: delete one comma
\end{point}

\reply{
    Whoops! Done.
}

### 2024-01-23 writeup.tex
bob note
new reviewer responses

Much of the work to 
understand the evolution of quantitative traits in sexual species
has focused on models in quantitative genetics
that describe evolution of trait values without explicitly representing underlying genetic variation,
but in a manner deriving from our understanding of how genetic variation works in practice.
Such models are therefore
interestingly intermediate between mechanistic models of genetics and purely phenomenological.
% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
% but tractable mathematical models of evolution for recombining species are elusive,
% at least with some results of when the model might be a good description of reality.
% For instance, a great deal of conceptual understanding and even quantitative predictions
% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
% despite the absence of a plausible model that would make this true (cite turelli and barton).
One concrete model of heritable trait evolution is the \emph{infinitesimal model},
which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
and a covariance independent of the parental trait values (but that depends on how they are related).
\citet{barton2017infinitesimal} give an overview of the complex history of the model, \revpoint{1}{12}
which is used in or adjacent to a good deal of theoretical work on quantitative genetics
\citep[e.g.,][]{fisher1918correlation,bulmer1971effect,lande1975maintenance}.
\citet{barton2017infinitesimal} focus on giving a formal proof of conditions
under which this would be a good approximation to reality,
a proof they later extended to include dominance in \citet{barton2022infinitesimal}.

As suggested by its name,
under the infinitesimal model the effect of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical studies of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many -- if not most -- lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps a trait is affected by mutations in a number of genomic regions,
and the distribution of effect sizes within each region is Normal, \revpoint{1}{13}
but with a standard deviation specific to the region that varies with ``proximity'' of that region to the trait
(e.g., regions that are further ``upstream'' have higher standard deviations).
Such mixtures can easily have large proportions of extreme values.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
and so the sum of $n$ copies, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
(On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$.) \revpoint{1}{11}

This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
in the infinitesimal model. \revpoint{1}{1}
In this paper, we make preliminary explorations in that direction:
Do the results of \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
The immediate answer to this will be ``no'' --
in fact, we will see that independence of offsprings' deviations and midparent values \revpoint{1}{2}
\emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense.
This independence is a key reason why the infinitesimal model is tractable,
and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
by ``trait-only'' models that do not explicitly represent underlying genetic variation.
Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
we then go on to analyze exactly these models.
(As motivation,
we have not shown that such models are \emph{not} good approximations to reality,
and it is interesting and perhaps informative to understand how the noise distribution affects results.)

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
only appear heritable through non-genetic means
such as systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
and $M$ is the number of loci. \revpoint{2}{2}
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|/2$.

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
Since $|X-1/2| = 1/2$, we'd like to know
if $\P\{|\eta|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |\eta| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta_{\ell}| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
(This weighting will make a difference, as frequencies and effect sizes are often correlated in practice.)
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
show that our estimates are unbiased but only accurate to within roughly $\pm 0.4$.
So, the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
a direct estimate of the underlying distribution of $\alpha$ values,
but it is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).

@@

Now we return to the theory.
We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$ as $M \to \infty$,
then with $\alpha \ge 2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
(The main result in \citet{barton2017infinitesimal} is that the trait values in generation $t$
are Gaussian conditioned on the traits in generation $t-1$,
but as discussed in that paper,
this implies that the traits of any set of individuals is multivariate Gaussian,
with covariance matrix determined by the pedigree.
Note, however, that this does \emph{not} imply that the distribution of trait values
across the population is Gaussian, as a collection of correlated Gaussians is not itself Gaussian.)

@@

We should now discuss what this counter-intuitive result says and does not say.
The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
and meiotic segregation is fair,
then independence of Mendelian sampling term from midparent value
\emph{implies} a joint Gaussian distribution for the two.
We have not assumed anything about the distribution of the trait in the population:
rather, we have showed that any model of this sort
that assumes independence of Mendelian sampling term and midparent value
must also, by logical necessity, have a Gaussian distribution for the two.
In other words, \revpoint{1}{7}
since models can be easily set up for which offspring minus midparent traits
are not Gaussian,
the implication of the proposition
is that for such models, independence of the Mendelian sampling term
is not likely a good assumption.
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between the effects
of alleles at such linked loci. \revpoint{2}{5}
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.
Finally, note that although they may not be independent,
the Mendelian sampling term $R_M$ and the midparent value $\bar Z_M$
\emph{are} always uncorrelated, since $Z_M$ and $V_M$ are exchangeable.

@@

As we discuss more later,
the presence of a large effect allele at a polymorphic locus \llabel{loci_polymorphic}
makes $R$ and $\bar Z$ dependent:
consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
Then in the limit $R$ depends strongly on parental trait values:
if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
then $R$ is close to $+5$ or $-5$ with equal probability.

% NOT HELPFUL
% Finally, a possible point of confusion is that \revpoint{1}{6}
% the infinitesimal model is often phrased as a statement about offspring trait distributions
% \emph{given} parental trait values,
% while our statement above regards parental trait values as random.
% However, the notions are equivalent:
% the statement that
% ``for each pair of parents the deviation of the offspring's trait from midparent
% has a distribution that does not depend on the parents' values''
% is essentially the definition of statistical independence
% of offspring's Mendelian sampling term and parental trait values.

@@

Although an infinitesimal model with non-Gaussian noise is incompatible with the usual additivity assumptions,
such models still exist (and might be useful);
next, we describe some of their properties.
As observed in \cite{turelli2017commentary},
there are at least three interpretations of what is called ``the infinitesimal model''. 
The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
assumes parental trait-independent Gaussian Mendelian sampling terms,
and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
this assumption is rather fragile --
a Gaussian distribution for the population only emerges under weak or Gaussian selection.
% However, the model has been used in influential work.
In principle, another factor that could produce a non-Gaussian population distribution
is if the Mendelian sampling terms are not Gaussian, as we consider here --
so, it is interesting to look for other situations where the population distribution is analytically tractable.

>>>>

Much of the work to 
understand the evolution of quantitative traits in sexual species
has focused on models in quantitative genetics
that describe evolution of trait values without explicitly representing underlying genetic variation,
but in a manner deriving from our understanding of how genetic variation works in practice.
Such models are therefore
interestingly intermediate between mechanistic models of genetics and purely phenomenological.
% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
% but tractable mathematical models of evolution for recombining species are elusive,
% at least with some results of when the model might be a good description of reality.
% For instance, a great deal of conceptual understanding and even quantitative predictions
% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
% despite the absence of a plausible model that would make this true (cite turelli and barton).
One concrete model of heritable trait evolution is the \emph{infinitesimal model},
which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
and a covariance independent of the parental trait values (but that depends on how they are related).
\citet{barton2017infinitesimal} give an overview of the complex history of the model,
which is used in or adjacent to a good deal of theoretical work on quantitative genetics
\citep[e.g.,][]{fisher1918correlation,bulmer1971effect,lande1975maintenance}.
\citet{barton2017infinitesimal} focus on giving a formal proof of conditions
under which this would be a good approximation to reality,
a proof they later extended to include dominance in \citet{barton2022infinitesimal}.

As suggested by its name,
under the infinitesimal model the effect of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical studies of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many -- if not most -- lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps a trait is affected by mutations in a number of genomic regions,
and the distribution of effect sizes within each region is Normal,
but with a standard deviation specific to the region that varies with ``proximity'' of that region to the trait
(e.g., regions that are further ``upstream'' have higher standard deviations).
Such mixtures can easily have large proportions of extreme values.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
On the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
and so the sum of $n$ copies, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
(On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$.)

This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
in the infinitesimal model.
In this paper, we make preliminary explorations in that direction:
Do the results of \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
The immediate answer to this will be ``no'' --
in fact, we will see that independence of offsprings' deviations and midparent values
\emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense.
This independence is a key reason why the infinitesimal model is tractable,
and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
by ``trait-only'' models that do not explicitly represent underlying genetic variation.
Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
we then go on to analyze exactly these models.
(As motivation,
we have not shown that such models are \emph{not} good approximations to reality,
and it is interesting and perhaps informative to understand how the noise distribution affects results.)

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, some important traits that have been claimed to be mostly genetically determined
only appear heritable through non-genetic means
such as systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
and $M$ is the number of loci.
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|/2$.

We do not consider the effects of new mutations.

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,  \revpoint{2}{1}
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
% $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
Since $|X-1/2| = 1/2$, we'd therefore like to know
if $\P\{|\eta|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |\eta| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta_{\ell}| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency,
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
(This weighting will make a difference, as frequencies and effect sizes are often correlated in practice.)
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.)
Examples are shown in Figure~\ref{fig:example_snps}.

@@

A possible concern is that tail exponents are difficult to estimate accurately,
but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
show that our estimates are unbiased but only accurate to within roughly $\pm 0.4$.
So, the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
a direct estimate of the underlying distribution of $\alpha$ values,
but it is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).

@@

Now we return to the theory.
We know from \citet{barton2017infinitesimal}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$ as $M \to \infty$,
then with $\alpha \ge 2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
(The main result in \citet{barton2017infinitesimal} is that the trait values in generation $t$
are Gaussian conditioned on the traits in generation $t-1$,
but as discussed in that paper,
this implies that the traits of any set of individuals is multivariate Gaussian,
with covariance matrix determined by the pedigree.
Note, however, that this does \emph{not} imply that the distribution of trait values
across the population is Gaussian, as a collection of correlated Gaussians is not itself Gaussian.)

@@

We should now discuss what this counter-intuitive result says and does not say.
The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
and meiotic segregation is fair,
then independence of Mendelian sampling term from midparent value
\emph{implies} a joint Gaussian distribution for the two.
We have not assumed anything about the distribution of the trait in the population:
rather, we have showed that any model of this sort
that assumes independence of Mendelian sampling term and midparent value
must also, by logical necessity, have a Gaussian distribution for the two.
In other words,
since models can be easily set up for which offspring minus midparent traits
are not Gaussian,
the implication of the proposition
is that for such models, independence of the Mendelian sampling term
is not likely a good assumption.
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between the effects
of alleles at such linked loci.
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.
Finally, note that although they may not be independent,
the Mendelian sampling term $R_M$ and the midparent value $\bar Z_M$
\emph{are} always uncorrelated, since $Z_M$ and $V_M$ are exchangeable.

@@

Finally, note that we follow \citet{barton2017infinitesimal} by assuming that loci are independent
(\ie no linkage disequilibrium) \revpoint{1}{1}
in the parental generation.
% (They are not necessarily identically distributed, however,
% so the Central Limit Theorem need not apply \emph{a priori}.)
One might then wonder if by relaxing this assumption
one could find a model in which the Mendelian sampling term was independent of midparent value
and yet had a non-Gaussian distribution.
To be interesting, such a model would need to maintain the nonindependence structure
that guarantees these properties,
which seems hard to arrange.

As we discuss more later,
the presence of a large effect allele at a polymorphic locus
makes $R$ and $\bar Z$ dependent:
consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
Then in the limit $R$ depends strongly on parental trait values:
if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
then $R$ is close to $+5$ or $-5$ with equal probability.

% NOT HELPFUL
% Finally, a possible point of confusion is that
% the infinitesimal model is often phrased as a statement about offspring trait distributions
% \emph{given} parental trait values,
% while our statement above regards parental trait values as random.
% However, the notions are equivalent:
% the statement that
% ``for each pair of parents the deviation of the offspring's trait from midparent
% has a distribution that does not depend on the parents' values''
% is essentially the definition of statistical independence
% of offspring's Mendelian sampling term and parental trait values.

@@

Although an infinitesimal model with non-Gaussian noise is incompatible with the usual additivity assumptions,
such models still exist (and might be useful);
next, we describe some of their properties.
As observed in \cite{turelli2017commentary},
there are at least three interpretations of what is called ``the infinitesimal model''. 
The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
assumes parental trait-independent Gaussian Mendelian sampling terms,
and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
As also discussed in \citet{turelli2017commentary},
this assumption is rather fragile --
a Gaussian distribution for the population only emerges under weak or Gaussian selection.
% However, the model has been used in influential work.
In principle, another factor that could produce a non-Gaussian population distribution
is if the Mendelian sampling terms are not Gaussian, as we consider here --
so, it is interesting to look for other situations where the population distribution is analytically tractable.

Leaving aside 
% -- as the Gaussian-population approximation frequently does --
the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if)
deterministic mutation-selection balance is possible in a model with non-Gaussian Mendelian sampling terms
that retains independence of these terms.
(Indeed, above we have put substantial restrictions on what models of inheritance
might lead to such models;
in particular,
we are \emph{not} working with the explicit sum-of-$\eta_\ell$ model of the previous sections.) \revpoint{1}{2}
We might also ask whether such a model
might still reasonably approximate an additive model with non-Gaussian effects,
and for the purpose of investigating this with simulations,
some calculations will be helpful.
